Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-08-27 manuscript.qmd
move prev. of midext to appendix for now

It is possible to provide our model with multiple different diagnostic modalities, each being characterized by different pairs of sensitivity and specificity. However, we instead chose to combine them into a single "consensus" diagnosis before parameter inference, as introduced in @sec-sec-data-consensus. We opted for this because the literature values of sensitivity and specificity [@de_bondt_detection_2007;@kyzas_18f-fluorodeoxyglucose_2008] of imaging modalities like MRI and CT do not plausibly match some of our observations: In the USZ cohort, 78% of OPSCC patients where diagnosed with ipsilateral LNL II involvement via diagnostic imaging. This is virtually impossible with sensitivities around 80% and specificities lower than 100%.

@@

### Prevalence of Midline Extension

Lastly, in @fig-model-prevalences-midext, we plot the prevalence of midline extension in the data versus our model's prediction. It is obvious the model cannot match the large spread between early and advanced T-category seen in the data. This is because to achieve that, it would need to increase the advanced T-category patient's prior distribution over diagnosis times and at the same time reduce the probability of the tumor to cross the midline during a time-step. But since the time-priors parameter is also coupled with the spread probabilities among the LNLs, the model does not have that freedom.

@@

>>>>

It is possible to provide our model with multiple different diagnostic modalities, each being characterized by different pairs of sensitivity and specificity. However, we instead chose to combine them into a single "consensus" diagnosis before parameter inference, as introduced in @sec-data-consensus. We opted for this because the literature values of sensitivity and specificity [@de_bondt_detection_2007;@kyzas_18f-fluorodeoxyglucose_2008] of imaging modalities like MRI and CT do not plausibly match some of our observations: In the USZ cohort, 78% of OPSCC patients where diagnosed with ipsilateral LNL II involvement via diagnostic imaging. This is virtually impossible with sensitivities around 80% and specificities lower than 100%.

@@

@@

# Prevalence of Midline Extension {.appendix}

<!-- I moved this here, because we think this is probably irrelevant or distracting w.r.t. the main findings of the paper. But I do think it is somewhat interesting. -->

In @fig-model-prevalences-midext, we plot the prevalence of midline extension in the data versus our model's prediction. It is obvious the model cannot match the large spread between early and advanced T-category seen in the data. This is because to achieve that, it would need to increase the advanced T-category patient's prior distribution over diagnosis times and at the same time reduce the probability of the tumor to cross the midline during a time-step. But since the time-priors parameter is also coupled with the spread probabilities among the LNLs, the model does not have that freedom.
